What to simulate? Inferring the direction for mental rotation

### 2014-01-30 man/cogsci-2014/cogsci-2014.tex
More work on paper

\textbf{Keywords:} 
\TODO{}
\end{abstract}

@@

In the classic experiment by \citeA{Shepard1971}, participants viewed
images of three-dimensional objects and had to determine whether the
images depicted the same object (which differed by a rotation) or two
separate objects (which differed by a reflection and a rotation). The
result was that people's response times had a strong linear
correlation with the minimum angle of rotation. The conclusion was
that people were solving this task by ``mentally rotating'' the
objects until they were congruent.

However, there are some questions regarding this process of problem
solving that are not explained by the alignment hypothesis: how do
people know the axis around which to rotate the objects? If the axis
is known, how do people know which direction to rotate the objects?
And finally, how do people know when to stop rotating and give a
response?

In this paper, we explore these questions through rational analysis
\cite{Marr:1983to,anderson90,Shepard:1987tt} and compare four models
of mental rotation. First, we discuss the previous literature on
mental imagery. Second, we outline computational- and
algorithmic-level analyses of the problem of mental rotation. Then, we
describe a behavioral experiment based on the classic mental rotation
studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of the
experiment with each of the models. We conclude with a discussion of
the strengths and weaknesses of each model, and lay out directions for
future work.

@@

Previous models of mental rotation have, for the most part, focused
mostly on the representation of mental images, rather than how people
decide \textit{which} mental images to
simulate. \citeA{Kosslyn:1977tv} propose a model of mental imagery's
visual buffer, but not the planning process of how it is used;
similarly, \citeA{Glasgow:1992tj} are mostly concerned with the
knowledge representation of imagery. Although \citeA{Anderson1978}
emphasizes the importance of considering both representation and
process, he dismisses the problem of determining the direction and
axis of rotation as merely a ``technical difficulty''.

The only model (of which the authors are aware) that has more
seriously attempted to address concerns regarding which simulation to
run is the parallel-process model of mental rotation by
\citeA{Funt:1983wn}. He determines the direction and axis of rotation
using the method-of-moments method; once this target rotation
computed, his model rotates one object until the target rotation is
reached, and then compares the two to see if they are congruent. This
is a somewhat unsatisfying explanation, however: why would it be
necessary to perform the rotation if the answer is already known?

@@

It may be useful to take a step back and try to analyze mental
rotation from using rational analysis
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we could say that the \textit{problem} is to determine
something about the spatial transformations that an object has
undergone. At the algorithmic level, we are constrained by the notion
that mental images must be transformed in an analog manner (or at
least in a way that is approximately analog), and that mental images
are time-consuming and effortful to generate. Thus, the \textit{goal}
is to make this determiniation while using the least amount of
resources. Due to the sequential constratiant, the minimum amount of
computation will coincide with the minimum angle of rotation. So, the
rational \textit{solution} will be something that only rotates through
the minimum angle.

motivate the active searching approach
\cite{Gureckis:2012gu,Markant:2012uu} \cite{Markant:2012uu,Nelson2007}

problem solving \cite{Hegarty2004, Schwartz1999}

symbolic vs simulation \cite{Schwartz:1996uy}

@@

In the mental rotation task, people are presented with pairs of
two-dimensional shapes similar to those used by \citeA{Cooper:1975wp};
an example stimulus is shown in Fig.~\ref{fig:stimuli}. The goal is to
determine whether the two images depict the same shape, or
mirror-image (flipped) shapes.

here we use simple 2d models as a starting point, but the general case
of solving this problem is very difficult, \TODO{cite shape matching
  literature}

@@

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$:
\begin{equation}
  p(\Xb\ \vert\ \Xa, \theta, h):= S(\Xb, f(\Xa, h, \theta)).
  \label{eq:likelihood}
\end{equation}
We do not know which vertices of $\Xb$ correspond to which vertices of
$\Xa$, so $S$ must marginalize over the set of possible mappings,
$\mathbb{M}$. For brevity, let $\mathbf{X}_m=\M\cdot{}f(\Xa, h,
\theta)$ where $\M\in\mathbb{M}$ is a permutation matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $\sigma_S=0.15$ is the standard deviation of the similarity.

Additionally, we assume that the observed shapes may only be
transformed by a small amount at a time, and each transformation is
costly in terms of computational resources. Specifically, if the
current mental image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, |\epsilon|) &\mbox{ rotate by $\epsilon$,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \mathcal{N}(0, \sigma_\epsilon)$ and
$\sigma_\epsilon=0.6$ is the standard deviation of the step size in
radians.

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
    \caption{\textbf{Model traces.} \TODO{}}
    \label{fig:model-traces}
  \end{center}
\end{figure*}

In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
similarity function, we must decide \textit{which} places to sample
and \textit{when} stop sampling. The models below differ in how they
make these decisions.

@@

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if computation produces a
valid rotation matrix; the $h$ that does produce a rotation matrix is
the correct hypothesis. This approach gives us the true value of
$\theta$, so Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio test,
where $\theta$ is set to the MLE value, rather than being
marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
  \label{eq:mle-lh-ratio}
\end{equation}

@@

We note that in the current formulation of the problem, choosing the
threshold is straightforward because we know both the exact geometry
of the shapes and that a linear transformation exists which will
perfectly align them. However, it is not always clear how to choose a
reasonable threshold \textit{a priori}, as the global optimum will
depend on many factors (such as shape complexity, dimensionality,
perceptual uncertainty, and whether the shapes are exactly
identical). Thus, it is unlikely that the \Th{} model will be robust
beyond this domain.

@@

While the previous few models have all focused on \textit{searching}
for the global maximum, the goal is really to just approximate
Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
rotations. Instead of perfoming a search, we could try maintaining a
probability distribution over our \textit{estimate} of
Eq.~\ref{eq:lh-ratio}, and then choose actions which are expected to
improve the accuracy of the estimate. This strategy has several
obvious benefits. First, it does not make assumptions about the shape
or scale of the similarity function. Second, by choosing to sample
places which are informative, this method implicitly minimizes the
amount of rotation, but because it is constrained to sequential
rotations, should still exhibit the linear effect of the angle.

@@

@@

We randomly generated 20 shapes of 5 or 6 vertices (e.g., the stimuli
in Fig.~\ref{fig:model-traces}). For each shape, we computed 20
``same'' and 20 ``flipped'' stimuli pairs, with $\theta$ spaced at
$20^\prime$ increments between 0 and 360 (with 0 and 180 degree
rotations repeated twice, in order to gather an equal number of
responses for each angle between 0 and 180). ``Same'' pairs were
created by rotating $\Xa$ by $\theta$; ``flipped'' pairs first
reflected $\Xa$ across the $y$-axis, then rotated by $\theta$.

We generated 5 shapes to be used in the practice block. Across all the
practice 10 stimuli, each shape and each angle (60, 120, 180, 240, or
300) was repeated twice (once ``flipped'' and once ``same'') such that
no shape was presented at the same angle twice. Additionally, we
generated a sixth shape to be included along with the instructions,
which had both a ``flipped'' and ``same'' version, each rotated to 320
degrees.

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time.pdf}
    \caption{\textbf{Response times.} \TODO{}}
    \label{fig:response-times}
  \end{center}
\end{figure*}

@@

All participants saw the same 10 practice trials (as described
above). There were a total of 800 experimental stimuli (20 shapes
$\times$ 20 angles $\times$ 2 same/flipped), which were split into 8
blocks of 100 across conditions in the following manenr. First,
stimuli were split into four blocks of 200 trials. Within each block,
each shape was repeated 10 times and each rotation was repeated 10
times (5 same, 5 flipped), such that across all blocks, each stimulus
appeared exactly once. Each block was then split in half. Each
participant then completed two half-blocks (not necessarily both from
the original full block).

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=0.85\textwidth]{../../figures/D/accuracy.pdf}
    \caption{\textbf{Accuracy.} \TODO{}}
    \label{fig:accuracy}
  \end{center}
\end{figure*}

@@

Out of the 247 participants, 200 (81\%) were included in our
analyses. Of the rest, we excluded 10 (4\%) because of an experimental
error, 6 (2.4\%) because they had already compelted a previous version
of the experiment, and 31 (12.6\%) because they failed a comprehension
check. A participant was marked as passing the comprehension check if
they answered at least 85\% of ``easy'' trials correctly. ``Easy''
trials were defined to be those for which the presented shapes
differed by a rotation of no more than 20 degrees. \TODO{mention
  excluding trials where people took a really long time to respond}

We ran model simulations for the same stimuli that participants
viewed. We ran 10 samples of each stimulus under each of the modesl,
with the exception of stimuli where $\theta=0$ or $\theta=180$, in
which case we ran double the number of samples.

@@

\begin{figure}[t]
  \centering
  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_accuracy.pdf}}%
  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_time.pdf}}
  \caption{\textbf{Effect of learning.} \TODO{}}
  \label{fig:learning}
\end{figure}

@@

There was a significant effect of learning both on response time
(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
though the effect on accuracy was not significant during the second
half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
vs. \ExpbTrialAccuracyCorr{} for the second
half). Fig.~\ref{fig:learning}

@@

The average response time across all stimuli was \ExpTime{} seconds,
and the average response time for an individual stimulus ranged from
\ExpTimeMin{} to \ExpTimeMax{} seconds.  As expected, the minimum
angle of rotation was significanty monotonically correlated with
average per-stimulus response times, both for flipped pairs
(\ExpThetaTimeCorrFlipped{}) and same pairs
(\ExpThetaTimeCorrSame{}). This was also the case for the number of
actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{} model
(\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
not the \Hc{} model
(\HcThetaTimeCorr{}). Fig.~\ref{fig:response-times}a

The number of actions taken by the \Oc{} model was the best predictor
of human response times, with a correlation of \ExpOcTimeCorr{}. The
\Th{} model was the next best (\ExpThTimeCorr{}), with the \Bq{} model
not far behind (\ExpBqTimeCorr{}). The \Hc{} model was not
significantly correlated with people (\ExpHcTimeCorr{}).

\TODO{flipped response times higher than same response times for
  smaller angles?}

@@

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} individual stimuli (out of 720) for which
people were not above chance. Of the models, the \Oc{} and \Th{}
models peformed perfectly, with 100\% accuracy. The \Bq{} model was
very accurate overall (\BqAccuracy{}), though there were
\BqNumChance{} stimuli for which it was not above chance. The \Hc{}
model was not particularly accurate, and although it did perform above
chance overall (\HcAccuracy{}), there were \HcNumChance{} individual
stimuli for which it was not above chance.

The minimum angle was also correlated with the average per-stimulus
accuracy: for flipped pairs, \ExpThetaAccuracyCorrFlipped{}, and for
same pairs, \ExpThetaAccuracyCorrSame{}. There was also a significant
effect for the \Hc{} model (\HcThetaAccuracyCorr{}) and the
\Bq{} model
(\BqThetaAccuracyCorr{}). Fig.~\ref{fig:accuracy}

There was a moderate correlation between the accuracy of the \Hc{}
model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
similar correlation, with \ExpBqAccuracyCorr{}.

>>>>

\textbf{Keywords:} mental rotation, computational model
\end{abstract}

@@

In the classic experiment by \citeA{Shepard1971}, participants viewed
images of three-dimensional objects and had to determine whether the
images depicted the same object (which differed by a rotation) or two
separate objects (which differed by a reflection and a rotation). The
result was that people's response times had a strong linear
correlation with the minimum angle of rotation, leading to the
conclusion that people solve this task by ``mentally rotating'' the
objects until they are congruent.  This conclusion leaves several
questions unanswered, however. How do people know the axis around
which to rotate the objects? If the axis is known, how do people know
which direction to rotate the objects?  And finally, how do people
know how long to rotate?

In this paper, we explore these questions through rational analysis
\cite{Marr:1983to,anderson90,Shepard:1987tt} and compare four models
of mental rotation. First, we discuss the previous literature on
mental imagery. Second, we outline computational- and
algorithmic-level analyses of the problem of mental rotation. Third,
we describe a behavioral experiment based on the classic mental
rotation studies \cite<e.g.,>{Cooper:1975wp}, and compare the results
of the experiment with each of the models. We conclude with a
discussion of the strengths and weaknesses of each model, and lay out
directions for future work.

@@

Previous models of mental rotation have, for the most part, focused on
the representation of mental images, rather than how people decide
\textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
propose a model of the mental imagery buffer, but not the planning
process of how it is used. Similarly, \citeA{Julstrom:1985va} and
\citeA{Glasgow:1992tj} are mostly concerned with modeling the
representational format underlying imagery. Although
\citeA{Anderson1978} emphasizes the importance of considering both
representation and process, he dismisses the problem of determining
the direction of rotation as a ``technical difficulty''.

The only model (of which the authors are aware) that has seriously
attempted to address the decision of \textit{what} to simulate is the
parallel-process model of mental rotation by \citeA{Funt:1983wn}. He
determines the direction, axis, and extent of rotation by computing
the objects' principal axes using moments of inertia. One object is
then rotated through the target rotation, and is checked against the
other object for congruency. However, this approach assumes that the
corresponding points on the two objects can be easily identified,
which is not necessarily the case.

@@

If people are not performing single-trajectory rotations, what might
they be doing? To answer this question, we perform a ``rational
analysis'' of the problem of mental rotation
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we could say that the \textit{problem} is to determine
something about the spatial transformations that an object has
undergone. At the algorithmic level, we are constrained by the notion
that mental images must be transformed in an analog manner (or in a
way that is at least approximately analog), and that mental images are
time-consuming and effortful to generate. Thus, the \textit{goal} is
to make this determination while performing a minimum amount of
computation (i.e., as few rotations as possible).

The original congruency hypothesis \cite{Shepard1971} actually is a
rational solution which acheives this goal, in the sense that the
smallest amount of computation coincides with rotating through the
minimum angle. However, it relies on assumptions of \textit{a priori}
knowledge (either of the minimum angle or the correspondence of points
between shapes). Noting that a rational solution need not necessarily
maintain a single trajectory of rotation, we explore an alternative
rational model, which--rather than assuming the angle of rotation
\textit{a priori}--engages in an ``active sampling'' strategy.

Active sampling is the idea that, when given information is
incomplete, people gather new information in a manner that increases
certainty about the problem space. This notion has recently gained
support in several areas of cognitive science
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, including other spatial
domains \cite{Juni:2011vo}. In the case of mental rotation, if people
do not know the minimum angle beforehand, then actively performing
rotations may be the best way to gather evidence about the similarity
between the observed shapes.

In the following section, we formalize this rational analysis, and
propose four models of mental rotation: one based on the original
hypothesis by \citeA{Shepard1971}; two which are extensions of the
first, but with relaxes assumptions; and one based on the active
sampling appraoach.

@@

In the mental rotation task, people are presented with pairs of
two-dimensional shapes similar to those used by \citeA{Cooper:1975wp};
example stimuli are shown in Fig.~\ref{fig:model-traces}. The goal is
to determine whether the two images depict the same shape, or
mirror-image (flipped) shapes.

@@

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$:
\begin{equation}
  p(\Xb\ \vert\ \Xa, \theta, h):= S(\Xb, f(\Xa, h, \theta)).
  \label{eq:likelihood}
\end{equation}
Example similarity curves are shown in Fig.~\ref{fig:model-traces}.

We do not know which vertices of $\Xb$ correspond to which vertices of
$\Xa$, so $S$ must marginalize over the set of possible mappings,
$\mathbb{M}$. For brevity, let $\mathbf{X}_m=\M\cdot{}f(\Xa, h,
\theta)$ where $\M\in\mathbb{M}$ is a permutation matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $\sigma_S=0.15$ is the standard deviation of the similarity.

Additionally, we assume that the observed shapes may only be
transformed by a small amount at a time, and each transformation is
costly in terms of computational resources. Specifically, if the
current mental image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, |\epsilon|) &\mbox{ rotate by $\epsilon$,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \mathcal{N}(0, \sigma_\epsilon)$ and
$\sigma_\epsilon$ is the standard deviation of the step size in
radians.

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
    \caption{\textbf{Example stimuli and model behavior.} On the left
      are two stimulus pairs, one ``same'' (top) and one ``flipped''
      (bottom), both with a rotation of $120^\prime$. The grey plots
      show the actions that each model took for these stimuli. The
      color of each point (action) indicates whether the model's
      current mental image was generated under the hypothesis that the
      objects are the same (red) or flipped (blue). The similarity
      functions for each of these hypotheses are shown on the right.}
    \label{fig:model-traces}
  \end{center}
\end{figure*}

In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
similarity function, we must decide \textit{which} places to sample
and \textit{when} stop sampling. The models below differ in how they
make these decisions (see Fig.~\ref{fig:model-traces} for examples of
their behavior).

@@

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
a valid rotation matrix; the $h$ that does is the correct
hypothesis. This approach gives us the true value of $\theta$, so
Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood ratio test, where
$\theta$ is set to the MLE value, rather than being marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
  \label{eq:mle-lh-ratio}
\end{equation}

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, it is not always clear how to choose a reasonable
threshold \textit{a priori}, as the global optimum will depend on many
factors (such as shape complexity, dimensionality, perceptual
uncertainty, and whether the shapes are exactly identical). Thus, it
is unlikely that the \Th{} model will be robust beyond this domain.

@@

While the previous few models have all focused on \textit{searching}
for the global maximum, the goal is really to just approximate
Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
rotations. Based on evidence that people use ``active sampling''
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a model that
actively monitors the usefulness of the rotations it is
performing. Instead of searching for a maximum, we maintain a
probability distribution over our \textit{estimate} of
Eq.~\ref{eq:lh-ratio}, and then sample actions which are expected to
improve that estimate.  This strategy has several benefits. First, it
does not make assumptions about the shape or scale of the similarity
function. Second, by choosing to sample places which are informative,
this method implicitly minimizes the amount of rotation; yet, because
it is constrained to sequential rotations, it should still exhibit the
linear effect of the angle.

@@

To evaluate the models described previously, we ran a behavioral
experiment based on classic mental rotation studies
\cite<e.g.>{Shepard1971, Cooper:1975wp}.

@@

We randomly generated 20 shapes of 5 or 6 vertices (e.g.,
Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
and 20 ``flipped'' stimuli pairs, with rotations ($\theta$) spaced at
$20^\prime$ increments between 0 and 360 (with 0 and 180 degree
rotations repeated twice, in order to gather an equal number of
responses for each angle between 0 and 180). ``Same'' pairs were
created by rotating $\Xa$ by $\theta$; ``flipped'' pairs first
reflected $\Xa$ across the $y$-axis, then rotated by $\theta$.

We generated 5 additional shapes to be used in a practice block of 10
trials. Across these trials, each shape and each angle (60, 120, 180,
240, or 300) was repeated twice (once ``flipped'' and once ``same'')
such that no shape was presented at the same angle twice. We also
generated a sixth shape to be included along with the instructions.
This shape had both a ``flipped'' and ``same'' version, each rotated
to 320 degrees.

@@

All participants saw the same 10 practice trials, as described
above. There were 720 unique experimental stimuli (20 shapes $\times$
18 angles $\times$ 2 same/flipped), though because stimuli with
rotations of $0^\prime$ or $180^\prime$ were repeated twice, there
were 800 total experimental stimuli. These stimuli were split across 8
conditions in the following manner: first, stimuli were split into
four blocks of 200 trials. Within each block, each shape was repeated
10 times and each rotation was repeated 10 times (5 same, 5 flipped),
such that across all blocks, each stimulus appeared exactly once. Each
block was then split in half. Participants completed two half-blocks
(not necessarily both from the original full block).

@@

@@

\subsection{Analysis}

@@

Out of the 247 participants, 200 (81\%) were included in our
analyses. Of the rest, we excluded 10 (4\%) because of an experimental
error, 6 (2.4\%) because they had already compelted a previous version
of the experiment, and 31 (12.6\%) because they failed a comprehension
check. A participant was marked as passing the comprehension check if
they answered at least 85\% of ``easy'' trials correctly. ``Easy''
trials were defined to be those for which the presented shapes
differed by a rotation of no more than 20 degrees. Additionally, we
excluded individual trials for which the response time exceeded 20
seconds.

For each model, we ran 50 samples for each of the same stimuli that
participants viewed (with the exception of stimuli with rotations of
$0^\prime$ or $180^\prime$, in which case we ran double the number of
samples). The step size parameter ($\sigma_\epsilon$) was
approximately fit to each of the models, resulting in
$\sigma_\epsilon=0.6$ for the \Bq{} model, and $\sigma_\epsilon=0.1$
for all other models. We also ran the models under two different
priors, $p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the
``unequal'' prior), though this only had an effect for the \Bq{}
model. \TODO{actually, did it affect HC?  check this}

\begin{figure}[t]
  \centering
  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
  \caption{\textbf{Effect of learning.} There was a clear effect of
    learning on participants' responses, especially in the first half
    of the experiment. Left: accuracy as a function of trial. Right:
    response time as a function of trial.}
  \label{fig:learning}
\end{figure}

@@

There was a significant effect of learning both on response time
(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
though the effect on accuracy was not significant during the second
half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
vs. \ExpbTrialAccuracyCorr{} for the second
half). Fig.~\ref{fig:learning} shows scatterplots illustrating these
trends.

@@

The average response time across all stimuli was \ExpTime{} seconds,
and the average response time for an individual stimulus ranged from
\ExpTimeMin{} to \ExpTimeMax{} seconds.  As expected, the minimum
angle of rotation was significanty monotonically correlated with
average per-stimulus response times, both for flipped pairs
(\ExpThetaTimeCorrFlipped{}) and same pairs
(\ExpThetaTimeCorrSame{}). This was also the case for the number of
actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{} model
(\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
not the \Hc{} model (\HcThetaTimeCorr{}). The top row of
Fig.~\ref{fig:response-time-accuracy} shows the overall effect of
rotation on response times.

The number of actions taken by the \Oc{} model was the best predictor
of human response times, with a correlation of \ExpOcTimeCorr{}. The
\Th{} model \Bq{} model with $p(h=0)=0.55$ were the next best (for
\Th{}, \ExpThTimeCorr{}; for \Bq{}, \ExpBqpTimeCorr{}). Not far behind
was the BQ model with equal prior (\ExpBqTimeCorr{}). The \Hc{} model
was barely correlated with people (\ExpHcTimeCorr{}).

Across all stimuli, participants' response times for ``flipped'' pairs
were significantly higher than those for ``same'' pairs
\TODO{confidence intervals} \TODO{flipped response times higher than
  same response times for smaller angles?}

@@

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} individual stimuli (out of 720) for which
people were not above chance. Of the models, the \Oc{} and \Th{}
models peformed perfectly, with 100\% accuracy. The \Bq{} model was
very accurate overall (equal prior: \BqAccuracy{}, unequal prior:
\BqpAccuracy{}). With the equal prior, there was \BqNumChance{}
stimulus for which it was not above chance; with the unequal prior,
there were \BqpNumChance{}. The \Hc{} model was not particularly
accurate, and although it did perform above chance overall
(\HcAccuracy{}), there were \HcNumChance{} individual stimuli for
which it was not above chance.

The minimum angle was also correlated with participants' average
per-stimulus accuracy: for flipped pairs,
\ExpThetaAccuracyCorrFlipped{}, and for same pairs,
\ExpThetaAccuracyCorrSame{}. There was also a significant effect for
the \Hc{} model (\HcThetaAccuracyCorr{}) and the \Bq{} model (equal
prior: \BqThetaAccuracyCorr{}, unequal prior:
\BqpThetaAccuracyCorr{}). The bottom row of
Fig.~\ref{fig:response-time-accuracy} shows the effect of angle on
accuracy for both participants and models.

There was a moderate correlation between the accuracy of the \Hc{}
model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
similar correlation (equal prior: \ExpBqAccuracyCorr{}, unequal prior:
\ExpBqpAccuracyCorr{}). Correlations for the other models were
undefined, because they had 100\% accuracy.

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
    \caption{\textbf{Response time and accuracy comparison.} Top:
      response time of correct responses as a function of the minimum
      angle of rotation. Bottom: accuracy as a function of the minumum
      angle of rotation. All error bars are 95\% confidence
      intervals.}
    \label{fig:response-time-accuracy}
  \end{center}
\end{figure*}
